---
title: "covid19analysis"
author: "Joseph"
date: "9/20/2020"
output: html_document
---
```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
                      fig.width=6.75, 
                      fig.height=6.75,
                      fig.align = "center",
                      error = TRUE, 
                      echo = TRUE
                      )
```


```{r load-libraries,echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(janitor)
library(png)
library(skimr)
library(vroom)
library(mosaic)
library(here)
library(rmarkdown)
library(ggExtra)
library(stringr)
library(dplyr)
library(reshape2)
library(tidyr)
library(forcats)

#remotes::install_github("kjhealy/socviz")
library(socviz)
```



As Covid-19 second wave bursts in Europes, let us try to use our analytical skills to uncover some trends in COVID-19 pandemic. 
We will try to determine whether prior medical conditions has any impact on death rate of COVID-19 patients, and whether there is a significant difference in death rate if an individual was admitted to Intensive Care Unit (ICU), depending on sex group and age. 


Let us revisit the [CDC Covid-19 Case Surveillance Data](https://data.cdc.gov/Case-Surveillance/COVID-19-Case-Surveillance-Public-Use-Data/vbim-akqf). 


```{r, cache=TRUE}
# URL link to CDC to download data
url <- "https://data.cdc.gov/api/views/vbim-akqf/rows.csv?accessType=DOWNLOAD"

covid_data <- vroom(url)%>%
  clean_names()
```



Let us start with Morbidity and COVID-19 death rate analysis. 

```{r, cache=TRUE}

death_morbidity <- covid_data %>% 
  
#We filter out all the "Missing" and "Unknown" values from the Death and Morbidity variables, we only keep "Male" and "Female" from the sexvariable and get eliminate any "Unknown" value in Age Group. 
          filter(medcond_yn == 'Yes' | medcond_yn == 'No', death_yn == 'Yes' | death_yn == 'No', 
                 sex == 'Male' | sex == 'Female', age_group != 'Unknown') %>%
  
#We select the variables of interest for our analysis.
          select(Age=age_group,Sex=sex,Death=death_yn,Morbidity=medcond_yn) %>% 
  
#We group our observation by Death, Age, Sex and Morbidity, the variables of interest for our graph. 
          group_by(Age ,Sex, Morbidity, Death) %>% 
          count()%>% 
  
#We modify our datafram to get "Yes" and "No" (Death) as Variable
          spread(key=Death, value=n) %>% 
  
#We summarise to get the death rate per Age group, Sex and Morbidity. We create an additional column to round those findings.
          summarise(death= (Yes/(Yes+No)*100),death_rate= round(death, digits = 1))

death_morbidity
```

Now that our data is ready, let us plot it in a bar chart to analyse the results. 

```{r, cache=TRUE}

#Let us add some labels to the graph. In order to add labels with the facet.grid, we need to create a function incorporating our label names. 
morbidity_label <- list('No' = "Without comorbidities", 'Yes' = "With comorbidities", 'Female'="Female", 'Male'="Male")
morbidity_labeller <- function(variable,value)
    {return(morbidity_label[value])}
  

#Let's move on to the plotting our findings, which will make it way easier to analyse our data.   
graph_death_morbidity <- death_morbidity %>% 
  
#Let us select the value necessary for the plot.
          ggplot(aes(y=death_rate, x=Age))+
  
#We need to plot a bar chart:  we don't need a legend but we add some color to the graph.
          geom_col(show.legend = FALSE, fill="#6B7CA4")+
  
# We want the x-axis to show the difference between man and women and y-axis to show the difference in comorbidity.
          facet_grid(Morbidity~Sex, scales = "fixed", labeller =(morbidity_labeller))+
  
#To make the graph more readable, let us put some labels and adjust their position
          geom_text(aes(label = death_rate, hjust = "left"), size = 2 , nudge_y = +1)+
  
#Let's put the scales in percentage, and swith the coordinates to have the death rates on the x axis. 
          scale_y_continuous(labels = scales::percent_format(scale = 1))+
          coord_flip()+
  
#Let's add a bit more style to our graph !
          theme_bw()+ 
  
#Let us add some titles, subtitles and additional elements of style
          labs(title= "Does past medical conditions really have an impact 
               on COVID mortality rate ?",
               subtitle = "Covid-19 Death rate by Age, Sex, and presence of Co-Morbidities\n",
               y="Death Rate (%)",
               x="Age Group",
               caption = "Source : CDC")+
          theme(plot.title = element_text(family = "Helvetica", face = "bold", size = (12)),
          axis.title = element_text(family = "Helvetica", size = (8)),
          axis.text = element_text(family = "Helvetica", size = (8))) %>% 
  guides(fill = guide_legend(override.aes = list(size=5), order = 2))
        
graph_death_morbidity

```

From this plot, it is evident that males have a much higher death rate for any category of age and morbidity than females. Also, we can see that the prior medical conditions have a critical impact on the COVID-19 death rate, for any age category. For instance, man aged over 80 years old with medical conditions have 67% death rate from COVID-19, compared to 29% death rate for cases without morbidity.
 
Let us now observe how the admittance to ICU has impact of the COVID-19 death rate per age category and sex. 

```{r, cache=TRUE}

death_ICU <- covid_data %>% 
  
#Similarly, we filter out all the "Missing" and "Unknown" values from the Death
#and ICU variables, we only keep "Male" and "Female" from the sex variable and get 
#eliminate any "Unknown" value in Age Group. 
          filter(icu_yn == 'Yes' | icu_yn == 'No', death_yn == 'Yes' | death_yn == 'No', 
                 sex == 'Male' | sex == 'Female', age_group != 'Unknown') %>%
  
#We select the variables of interest for our analysis.
          select(Age=age_group,Sex=sex,Death=death_yn,ICU=icu_yn) %>% 
  
#We group our observation by Death, Age, Sex and ICU, the variables of interest for our graph. 
          group_by(Age ,Sex, ICU, Death) %>% 
          count()%>% 
  
#We modify our dataframe to get "Yes" and "No" (Death) as Variable
          spread(key=Death, value=n) %>% 
  
#We summarise to get the death rate per Age group, Sex and Morbidity.
          summarise(death2= (Yes/(Yes+No)*100), death_rate2= round(death2, digits = 1))

death_ICU
```

Now that the data is ready, let us plot it to make it more readable. 

```{r, cache=TRUE}


#Let us add some labels to the graph. In order to add labels with the facet.grid, we need to create a function incorporating our label names. 
icu_label <- list('No' = "Not admitted to ICU", 'Yes' = "Admitted to ICU", 'Female'="Female", 'Male'="Male")
icu_labeller <- function(variable,value)
    {return(icu_label[value])}
  

#Let's move on to the plotting our findings, which will make it way easier to analyse our data.   
graph_death_icu <- death_ICU %>% 
  
#Let us select the value necessary for the plot.
          ggplot(aes(y=death_rate2, x=Age))+
  
#We need to plot a bar chart:  we don't need a legend but we add some color to the graph.
          geom_col(show.legend = FALSE, fill="#EF8E7B")+
  
# We want the x-axis to show the difference between man and women and y-axis 
# to show the difference in comorbidity.
          facet_grid(ICU~Sex, scales = "fixed", labeller = icu_labeller)+
  
#To make the graph more readable, let us put some labels and adjust their position
          geom_text(aes(label = death_rate2, hjust = "left"), size = 2 , nudge_y = +1)+
  
#Let's put the scales in percentage, and swith the coordinates to have the 
#death rates on the x axis. 
          scale_y_continuous(labels = scales::percent_format(scale = 1))+
          coord_flip()+
  
#Let's add a bit more style to our graph !
          theme_bw()+ 
  
#Let us add some titles, subtitles and additional elements of style
          labs(title= 
  "Does past medical conditions really have an impact on COVID mortality rate ?",
               subtitle = "Covid-19 Death rate by Age, Sex, and presence of Co-Morbidities\n",
               y="Death Rate (%)",
               x="Age Group",
               caption = "Source : CDC")+
          theme(plot.title = element_text(family = "Helvetica", face = "bold", size = (12)),
          axis.title = element_text(family = "Helvetica", size = (8)),
          axis.text = element_text(family = "Helvetica", size = (8)))
        
graph_death_icu
          
```
That looks great! From this graph, it appears that admittance to Intensive Care Unit has a great impact on the COVID-19 death rate. Similarly to the previous graph, males are also more impacted by the pandemic than females for almost any age category. 



